---
title: "Multinatinal Bank Customer Churn"
output: 
  html_document:
    theme: "spacelab"
    highlight: kate
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 6
    toc_float: yes
    collapsed: false
    smooth_scroll: false
  pdf_document:
    toc: yes
    toc_depth: '3'

---

```{r  include=FALSE,echo=FALSE}
library(knitr)
library(ezids)
loadPkg("Hmisc") 
library(dplyr)
loadPkg("MASS")
library(ggplot2)
knitr::opts_chunk$set(warning = F, message = F,echo=FALSE)
options(scientific=T, digits = 3) 
```

# Chapter 1: Introduction

Customer churn, also known as customer turnover, occurs when a customer chooses to leave or unsubscribe from a service for any reason. In the context of this project, we are looking at customer churn in banking. Being able to predict customer churn is important because we want to retain as many customers as we can. 

Of course you want to keep customers, but it also is cheaper to retain your customers, than having to spend money to acquire new customers to replace the ones that left. Lower customer acquisition costs equal more profits. By identifying customers at risk of churn, we are able to identify which efforts in order to maximize their likelihood of staying.  

We chose to analyze a multinational bank’s churn for several reasons. As an international team, we were interested in topics that spanned multiple countries, with a particular focus on either business or environment, our personal interests. When we were pitching topics, nearly every suggestion was geared towards sales or customer experience, or global environmental issues.

After ensuring that the data sets that interested us had enough observations to follow the project’s 3,000 row minimum, we then identified the topics with the highest number of useful variables. There were several data sets, which, while interesting, did not yield enough useful variables. Rather, they included variables like unique product, user, or review identifiers, which were not relevant to the type of analysis we were performing. 

From our short-list, the Bank Customer Churn Data Set had the highest number of useful variables, 11 out of the 12, as opposed to 4-6 useful variables for our runner-up options. We made the decision to go with our chosen data set because we figured that a larger number of variables (and data in general) is a closer reflection of the multitude of factors that affect a customer’s decision to churn. Additionally, we do not know at first glance which of the variables will end up being statistically significant, however, by having a higher number of variables, we are considering the potential effects of a wider range of variables, which may be useful to our model after we perform EDA.

After finding a data set that piqued our personal interests, met the project requirements, and had a larger range of variables than the other top choices, we were satisfied with our choice and able to proceed to the next stage.

In our research, we aim, through EDA, to identify the variables which have the highest correlation with churn, across three countries serviced by a multinational bank. The variables identified can then be used to inform banks of what customer activity trends and attributes should be monitored so the bank can make efforts to retain the customers before the account has been closed.


# Chapter 2: Description of the Data

This study is based on a data set that contains 10,000 observations of bank customers with 12 variables. (See below for a readout of the dataset’s structure and variable names.) Variable descriptions are as follows and we choose the data set from [Kaggle](https://www.kaggle.com/code/tamunoyedarego/abc-multinational-bank-customer-analysis).


```{r Churn_data_file, include=T,echo=FALSE,results='apois'}
churn_data<-data.frame(read.csv("Bank_Customer_Churn_Prediction.csv",header=TRUE)) 
str(churn_data) 
```

1. **Customer ID** - The Unique ID of each individual customer
2. **Credit Score** - A number depicting the customer's creditworthiness
3. **Country** - The country the customer banks from
4. **Gender** - The gender the customer identifies with
5. **Age** - The customers age
6. **Tenure** - Indicates how length in years the customer has been with the bank
7. **Balance** - The amount currently available in the customer's account
8. **Products Number** - The number of products purchased by the customer through the bank
9. **Credit Card** - Indicates the customer has a credit card
10. **Active Member** - Indicates if the customer is an active or inactive
11. **Estimated Salary** - Bank Estimation of the income of the customer
12. **Churn** - Indicator of if the customer has left the bank or not

A limitation of the data set is that it is imbalanced, as the customer country is not evenly distributed. There are three countries, however half of all customers are based in France, and the remaining two countries are each home to a quarter of all customers. When creating our model, we will need to take this into account and plan to handle this imbalanced data.

A second limitation is that some of the variables in the data set are poorly defined. One such variable is ‘active_member’, which does not operationally define what it means to be an active member (ie. has not logged in in X months, has had no deposits in time period X). Another variable which does not provide enough information about its definition is products_number. While it does give the number of products an account holder has purchased, we do not know which products, and in which combination they are used by the customers. While we can draw conclusions based on the number of products, we cannot make further inferences about the customer's behavior based on the type of product or effects relating to combinations of products.


# Chapter 3: Independent Variables EDA:
In preparation for exploratory data analysis, we took several steps to clean the data. We immediately dropped the customer_id variable, as we do not need the account holder’s unique identifier for our purpose. After dropping customer_id our dataset had 11 variables:

```{r Remove cutomer_id from the dataframe, include=TRUE, echo=FALSE,result='asis'}
churn_data <- churn_data[-c(1)] # removing column 1(customer _id) from data set
str(churn_data)
```

Next, we checked for duplicate records in the data set,we found no duplicates in the data set and also checked for null variables, and 0 were found.We then converted the following variables into categorical variables:credit_card, active_member, churn, gender, tenure, product number, age.We also converted Boolean values (0,1) into character format for Credit_card (credit_card,no-credit card), active_members (Active,In Active) and churn (Churned, Retained) variables in the data set, in order to understand each variable in the data set clearly during plotting and analysis.


```{r  echo=FALSE,result='asis'}
library(knitr)
churn_data$active_member[churn_data$active_member== 1]<-"Active"
churn_data$active_member[churn_data$active_member== 0]<-"In Active"

churn_data$credit_card[churn_data$credit_card== 1]<-"Credit Card"
churn_data$credit_card[churn_data$credit_card== 0]<-"No-Credit Card"

churn_data$churn[churn_data$churn== 1]<-"Churned"
churn_data$churn[churn_data$churn== 0]<-"Retained"
kable(head(churn_data[,c(8,9,11)]))
```

Finally, we checked our continuous variables for outliers using the outlierKD function.

```{r customer_credit_score,include=TRUE, echo=FALSE,result='asis'}
credit_score_clean <- outlierKD2(churn_data, credit_score, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
age_clean <- outlierKD2(churn_data, age, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
```

By using the outlierKD function we can observe that outliers were found only in `age` and `credit_score` variables (age: 3.7% and credit_score: 0.2%), so we decided to remove these two variables using the outlierKD function.

## Customer Analysis :Slicing the data for an Overview

In this section we are going to analyze each variable in the Bank churn dataset by using plots along with finding calculated mean,SD and percentages for each variable.

### What is the average credit score of the customers?

The average credit score of the customer is 650.529,most of the customers having credit score fall between 600 to 700 and standard deviation is 96.953.The below histogram shows the range of credit scores.

```{r Credit_score}
library(ggplot2)
ggplot(data=churn_data, aes(x=credit_score)) + 
  geom_histogram(col="black",bins=30,
                 fill="dark orange", 
                  alpha =.7) + # opacity
  labs(title="`ggplot`") +
  labs(title="Customer's Credit Scores",x="Customer Credit Score", y="Customer Count") 
```


### which are the different countries in which a customer holds a bank account?

The customers are grouped by the countries in which they have their accounts.As we see from the plot France has more than 50% of customer accounts which is the highest among all other countries with Germany and Spain sharing equal percentages.Below bar plots represents number of customers in three different countries. 

```{r Country}
ggplot(data=churn_data,aes(x=country,fill=country))+
  geom_bar(col="black")+
  scale_fill_brewer(palette="Reds") +
  labs(title = "Bank Customer vs Country",x="Country",y="Customer Count")+
  theme_minimal()
```

### What is the average age of the customer?

The majority of the bank customer's fall below the age of 50 with average age of 39 and with standard deviation of 10.5.

```{r Age}
ggplot(churn_data, aes(x = age,fill=cut(age,100))) + 
  geom_histogram(show.legend = FALSE,col="black",bins=30)+
  scale_fill_discrete(h=c(240,10),c=120,l=70)+
  theme_minimal()+
  labs(x=" Customer Age",y=" Customer Count")+
  ggtitle("Customer ages")
```

### How many years does the customer have a relationship with the bank?

We can see from the histogram plotted below that most of the customers have been with the bank for more than a year.

```{r tenure}
library(ggplot2)
ggplot(data=churn_data, aes(x=tenure)) + 
  geom_histogram(col="black",bins=11,
                 fill="Yellow", 
                  alpha =.7) + 
  labs(title="Years spent as customer",x="Number of years", y="Customer Count") 
```

### What is the percentage of Male and Female customers who hold an account in the bank?

From the pie chart below we can see that the bank's customers are predominantly male which make up 55% of the customer base with the females making up the remaining 45%.

```{r Gender}
df <- churn_data %>% 
  group_by(gender) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))


ggplot(df, aes(x = "", y = perc, fill =gender )) +
  geom_col(color="black") +
  geom_text(aes(label = labels),color = c("black", "black"),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y")+
  labs(title="Percentage of Male and Female customers in bank",x="", y="") 
``` 

### What is the percentage of active and Inactive account holders in the bank?

Huge percentage of customers are being inactive,it is reported that 48.5% of customers are being inactive.

```{r Active_Memebers}
df <- churn_data %>% 
  group_by(active_member) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

ggplot(df, aes(x = "", y = perc, fill = active_member)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "white"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_grey()+
  labs(title="Percentage of active members in bank",x="", y="") 
``` 

### What are the different types of services/Products purchased by customers provided by the bank for better usage of baking services?
Most of the customers use product 1 which is 50% and product 4 is the least used with 0.6%.

```{r Products}
df <- churn_data %>% 
  group_by(products_number) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))


ggplot(df, aes(x = "", y = perc, fill = products_number)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("white", "white","white","white"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  labs(title="Percentage of  different products used by customers ",x="", y="") 
  
``` 

### What percentage of customers make use of credit cards?
Predominantly 71% of the Bank customers use credit_card and only 29% do not make use of it.

```{r Credit_Card}
df <- churn_data %>% 
  group_by(credit_card) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))


ggplot(df, aes(x = "", y = perc, fill = credit_card)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "black"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_brewer(palette="Greens")+
  labs(title="Percentage of customer use credit card",x="", y="") 
```  

### Currently what percentage of customers are churned from the bank?
The bank managed to retain 80% of their customers with the remaining 20% where churned out.

```{r Customer_churn}
df <- churn_data %>% 
  group_by(churn) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))


ggplot(df, aes(x = "", y = perc, fill = churn)) +
  geom_col(color="black") +
  geom_label(aes(label = labels),color = c("black", "black"),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")+
  scale_fill_brewer(palette="Purples")+
  labs(title="Percentage of customer retained",x="", y="") 
``` 

So,now the big question was what factors affected the 20% churn rate in the multinational bank across different countries.We will do  statistical analysis to find the cause for the churn rate by framing a few possible SMART questions.


# Chapter 4 : Independent variables EDA:Bar Plots,Histograms,Chi Square test,Correlation plot.

### SMART question: Does the churn rate depend on Gender of the account holders?

Chi Square test for independence
H0 : Customers' gender and churn are independent.
H1 : Customer’s gender and churn are not independent.

```{r Chisq test of independence between Gender and customer churn rate}
#Contingency table
contab = table(churn_data$churn, churn_data$gender)
xkabledply(contab, title="Contingency table for Gender(Male, Female) vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab)
chitests
```

We have $\chi^2$  value of the test from `chitests$statistic` = `r chitests$statistic`,while extracting the p-value from `chitests$p.value` = `r chitests$p.value` which is less than significance level of 0.05 ,Hence we reject the null hypothesis and we conclude that the churn rate dependent on gender of the customer.

```{r Gender_churns}
ggplot(data=churn_data,aes(x=gender,fill=churn))+
  geom_bar(col="black")+
   scale_fill_manual('Position', values=c('red', 'lightblue'))+
  labs(title = "Gender vs Churn",x="Gender",y="Chrun")+
  theme_minimal()
``` 

Here, we are comparing the number of male and female customers who are about to churn. Out of 54.6% male customers only 16.5% customers are churned. whereas, on the other hand 25.07% female customers are churned of 45.4%. From the plot it is observed that majority of the churned customers were female despite the total population of the bank being predominantly male. It is clear that there is a sufficiently high enough correlation to suggest that gender and customer attrition are dependent variables.

### SMART question:Does the churn rate depend on the account balance of the customer?



### SMART question: Does the account holder's age groupo affect churn?

Correlation test between age and churn

### SMART question: Does churn depend on status of active accounts?

Chi Square test for independence
H0 : Customers' active status and churn are independent.
H1 : Customer’s active status and churn are not independent.

```{r Chisq test of independence between Active members and customer churn rate}
#Contingency table
contab_activemember = table(churn_data$churn, churn_data$active_member)
xkabledply(contab_activemember, title="Contingency table for Account Status(Active, In Active) vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab_activemember)
chitests
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis and find that the churn rate is reliant on the customer's active status.

```{r Active_members_churns}
ggplot(data=churn_data,aes(x=active_member,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'pink1')) +
  labs(title = "Active Members vs Churn",x="Active Status of Account Holders",y="Churn")+
  theme_minimal()
```  

From the boxplot, we are comparing the number of Active and Inactive customers who are about to churn. A large number of consumers are inactive; it has been stated that 48.5% of clients are inactive. Out of 51.51% Active customers only 14.26% customers are churned. whereas, on the other hand 26.85% Inactive customers are churned of 48.49%. It is clear that there is a sufficiently high enough correlation to suggest that active status and customer attrition are not independent variables. As a result, we may conclude that the majority of Inactive clients are about to churn.

### SMART question: Does churn depend on different banking product and services provided?

Chi Square test for independence
H0 : Banking products,services and churn are independent.
H1 : Banking products,services and churn are not independent.

```{r Chisq test of independence between Banking product and services and customer churn rate}
#Contingency table
contab_productnumber = table(churn_data$churn, churn_data$products_number)
xkabledply(contab_productnumber, title="Contingency table for Banking Services vs Churn (0 or 1)")

#Chisquare Test of independency
chitests = chisq.test(contab_productnumber)
chitests
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis and find that the churn rate is reliant on the customer's banking product and services.

```{r Bank_Services_churns}
ggplot(data=churn_data,aes(x=products_number,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lavender')) +
  labs(title = "Banking Product and Services vs Churn",x="Banking Product and Services",y="Chrun")+
  theme_minimal()
```  

We can notice that most of the customers with lower and higher products i.e) 1, 3 and 4 are about to churned when compared to customers with two products.

### SMART question: Does churn depend on credit score?

Correlation test between credit score and churn

### SMART question: Does a uers having a credit card with the bank affect predisposition to churn?

Chi Square test for independence
H0 :Customer having credit card and churn are independent.
H1 : Customer having credit card  and churn are not independent.

```{r Chisq test of independence between credit card and customer churn rate}
churn_data[, 10:11][churn_data[, 10:11] == 1] <- 'Churn'
churn_data[, 10:11][churn_data[, 10:11] == 0] <- 'No Churn'

cc_churn <- table(churn_data$churn, churn_data$credit_card)
xkabledply(cc_churn, title="Contingency table")

chitestccchurn <- chisq.test(cc_churn)
chitestccchurn
```

```{r Credit_Card_churns}
ggplot(data=churn_data,aes(x=credit_card,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lightblue')) +
  labs(title = "Customer Churn in Customers With and Without Credit Cards",x="Credit Card",y="Churn")+
  theme_minimal()
```

Credit Card vs Churn
P-value of 0.5 is much higher that the significance level (0.05 for df=1). Thus, we reject the null hypothesis, as there is a 0.5 or 50% chance of these results occurring by chance. We can conclude that there is no relationship between whether a user has a credit card, and churn.

### SMART question: Does churn rate depend on which country the user banks in?

Chi Square test for independence
H0 : Country card and churn are independent.
H1 : Country and churn are not independent.

```{r Chisq test of independence between country and customer churn rate}
#Contingency table
contable = table(churn_data$country, churn_data$churn)
xkabledply(contable, title="Contingency table for Country (Customer belongs to) vs Churn (0 or 1)")

#Chisquare Test of independency
chitest <- chisq.test(contable)
chitest
```

We have the $\chi^2$ value of the test from `chitests$statistic` = `r chitests$statistic`, while the p-value is less than the significance level of 0.05 from `chitests$p.value` = `r chitests$p.value`. As a result, we reject the null hypothesis and find that the churn rate is reliant on the customer's regions.

```{r Country_churns}
ggplot(data=churn_data,aes(x=country,fill=churn))+
  geom_bar()+
  scale_fill_manual('Position', values=c('red', 'lavenderblush2')) +
  labs(title = "Customer Churn in Different Countries",x="Country",y="Churn")+
  theme_minimal()
``` 

The bar plots represents number of customers in three different countries. The customers are grouped by the countries in which they have their bank accounts. As we see from the plot France has more than 50% of customer accounts which is the highest among all other countries with Germany and Spain sharing equal percentages.

In terms of churn rate, the percentage of customers churned in the France about 16.15% on this as opposed to the Spain at 16.67%. On the other hand Germany has the highest churned (around 32.44%) was over doubled that of France and Spain. Hence most of the customers from Germany are about to churned when compared to customers from France and Spain.

# Chapter 5: Tests and Results
In our EDA, we utilized Chi-squared of independence, One Sample T-test, Two Sample T-test, Z-Test, and correlation matrix (corrplot).

# Chapter 6: Conclusions

# Chapter 7: Bibliography

